Service Oriented ARC Storage; \\An Overview and Design

### 2008-12-04 doc/storage_system/ARCStorage_src/ARCStorage.tex
Implemented comments from Bjarte

\begin{document}
\pagestyle{headings} 
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$  \\ $^{*}$Scientific
  Computing Division, IT Department, Uppsala University, Sweden
  \\ $^{\dagger}$Center for Information Technology, University of Oslo, Norway
  \\ $^{\ddagger}$Department of Physics,  University of Oslo, Norway
  \\  $^{\mathsection}$Department of Physics and Astronomy, Uppsala University,
  Sweden \\ $^{\mathparagraph}$National Information Infrastructure Development
  Institute, Hungary\\ \\  j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

elements in the order of several hundreds of terabytes in a single

@@

Bigtable. Chubby has several features similar to the distributed A-Hash service,

@@

Bigtable is neither free, open-source, nor available to the public.

@@

The Storage Resource Broker (SRB) \cite{earlySRB,SRB} provides a flexible data grid management

@@

offered by our Gateway service. However, being built as a middleware on top of

@@

The next generation Advanced Resource Connector (ARC) Grid middleware \cite{NorduGridsite} \cite{KnowARCsite} consists of a set of pluggable components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services are running inside a container called the Hosting Environment
Daemon (HED). There are four kinds of pluggable components having well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors and
Policy Decision Components are responsible for the security model
within the system.

@@

In the default implementation the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC data management system in the foreseen
future is restricted to provide support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

@@

As is the case for all openly accessible web services, the security
model is of crucial importance for the ARC Storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} takes care of the
  internal communication between services. There is a lot of
  communication between the services of the ARC Storage. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians, the Librarians talk with
  the A-Hash. If any of these services get compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested the Shepherd will provide a one-time
  Transfer URL (TURL) which the client can connect to. In the most
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of compromitation is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, offering a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with regular X.509 authentication.

@@

The services were distributed as shown in figure \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs on UPPMAX.
\item In total we are running five Shepherds: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

@@

The following test where considered:

\begin{itemize}
\item Test 1: The distribution after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading
\item Test 3: Distribution after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the shepherd becomes offline.   
\end{itemize}

@@

Figure \ref{fig:FileDistribution} illustrate two different scenarios,
corresponding to Test 2 and Test 3: The orange bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green bars show distribution with the same kind of
uploading, but with the clients requesting three replicas each. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) is not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in figure \ref{fig:FileDistribution}.

@@

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llll}
\hline \hline
&\small{Before (MB)}&\small{During (MB)}&\small{After (MB)}\\
\hline
\small{UPPMAX}&4888&8799&8798\\
\small{USIT-1}&7820&9778&6843\\
\small{USIT-2}&6843&-&4888\\
\small{USIT-3}&7820&9774&8798\\
\small{NIIF}&3320&2344&1367\\
\hline
\small{Sum}&30691&30693&30694\\
\hline
\end{tabular}
\caption{Storage load before, during and after a Shepherd outage.}
\label{table:self healing}  
\end{table}

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking out one Shepherd from the system, and later reinsert it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished; and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted. Interestingly, the storage load
automatically evens out when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. Interestingly, the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaniously
starts uploading replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

>>>>

\begin{document}
\pagestyle{headings} 
%% \title{Service Oriented ARC Storage; \\An Overview and Design}
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$  \\ $^{*}$Scientific
  Computing Division, IT Department, Uppsala University, Sweden
  \\ $^{\dagger}$Center for Information Technology, University of Oslo, Norway
  \\ $^{\ddagger}$Department of Physics,  University of Oslo, Norway
  \\  $^{\mathsection}$Department of Physics and Astronomy, Uppsala University,
  Sweden \\ $^{\mathparagraph}$National Information Infrastructure Development
  Institute, Hungary\\ \\  j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

elements to collect several hundreds of terabytes in a single

@@

Bigtable. Chubby has several features similar to our A-Hash service (see Section \ref{Hash Algorithm}),

@@

Bigtable is neither free, open-source nor available to the public.

@@

the Storage Resource Broker (SRB) \cite{earlySRB,SRB} provides a flexible data grid management

@@

offered by our Gateway service (see Section \ref{Gateways}). However, being built as a middleware on top of

@@

The next generation Advanced Resource Connector (ARC) Grid middleware \cite{NorduGridsite} \cite{KnowARCsite} consists of a set of pluggable components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services are running inside a container called the Hosting Environment
Daemon (HED) and there are four kinds of pluggable components having well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors and
Policy Decision Components are responsible for the security model
within the system.

@@

In the default implementation the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC storage system in the foreseen
future is restricted to provide support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

@@

As is the case for all openly accessible web services, the security
model is of crucial importance for the ARC Storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} takes care of the
  internal communication between services. There is a lot of
  communication between the services of the ARC Storage. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians and the Librarians talk with
  the A-Hash. If any of these services get compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested the Shepherd will provide a one-time
  Transfer URL (TURL) which the client can connect to. In the most
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of compromitation is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, offering a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with regular X.509 authentication.

@@

The services were distributed as shown in figure \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs on UPPMAX.
\item In total five Shepherds were used for the tests: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

@@

The following test where carried out:

\begin{itemize}
\item Test 1: The distribution of data after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution of files after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading
\item Test 3: The distribution of data after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the shepherd becomes offline.   
\end{itemize}

@@

Figure \ref{fig:FileDistribution} illustrate two test results,
corresponding to Test 2 and Test 3: The orange bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green bars show distribution with the same kind of
uploading, but with the clients requesting three replicas for each file. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) is not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in figure \ref{fig:FileDistribution}.

@@

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llll}
\hline \hline
&\small{Before (MB)}&\small{During (MB)}&\small{After (MB)}\\
\hline
\small{UPPMAX}&4888&8799&8798\\
\small{USIT-1}&7820&9778&6843\\
\small{USIT-2}&6843&-&4888\\
\small{USIT-3}&7820&9774&8798\\
\small{NIIF}&3320&2344&1367\\
\hline
\small{Sum}&30691&30695&30694\\
\hline
\end{tabular}
\caption{Storage load before, during and after a Shepherd outage.}
\label{table:self healing}  
\end{table}

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking out one Shepherd from the system, and later reinsert it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished; and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted and that the storage load evens out 
again when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. We also see that the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaniously
starts uploading replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.
